### 2023-06-24 Draft.tex
Written types of Filtering

\[        
    \pmb{K} = \sum_{i}{\sigma_i \cdot u_i \cdot v_i^{\pmb{T}}} 
\]
or more generally
\begin{align*}
        \pmb{K} =& \textbf{U}_{m\times p} \pmb{\textstyle\sum}_{p\times p} \textbf{V}_{p\times n}^{\pmb{T}} \\
        \pmb{K} =& \begin{bmatrix}
            u_0 & \hdots & u_{p-1}
        \end{bmatrix}
        \begin{bmatrix}
            \sigma_0&&\\
            &\ddots&\\
            &&\sigma_{p-1}
        \end{bmatrix}
        \begin{bmatrix}
            v_0 \\
            \vdots \\
            v_{p-1}
        \end{bmatrix}
\end{align*}
Note that it requires the number of rows and columns must be the same, we can also alleviate this by putting 0 appropriately,
to allow the kernel to be a square matrix.Then the first singular value $\sigma_0$ is checked. If it is the only one that is non zero,
and the rest is zero, then the kernel is separable. As an example, the sobel filter, which is used for edge detection, 
can be separated into two separate operations. In this case, this is for the horizontal derivative approximation.
\[
    \frac{1}{8}
    \begin{bmatrix}
        -1 & 0 & 1 \\
        -2 & 0 & 2 \\
        -1 & 0 & 1
    \end{bmatrix}
    =
    \frac{1}{8}
    \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
    \begin{bmatrix} 1 \\ 2 \\ 1 \end{bmatrix}
\]
The normalization of the operators are easily calculated by dividing each operator with the absolute sum of the entries.
\[  
    \begin{bmatrix} -1 & 0 & 1 \end{bmatrix} \Rightarrow \vert -1 \vert + \vert 0 \vert + \vert 1 \vert = 2 \Rightarrow \frac{1}{2}
\]
Then the normalized operator is
\[
    \frac{1}{2} \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
\]
it applies to the vertical operator as well
\[
    \frac{1}{4} \begin{bmatrix} 1 \\ 2 \\ 1 \end{bmatrix}
\]
The following are more examples of common separable filtering.

@@

>>>>

\[        
    \pmb{K} = \sum_{i}{\sigma_i \cdot u_i \cdot v_i^{\pmb{T}}} 
\]
or more generally
\begin{align*}
        \pmb{K} =& \textbf{U}_{m\times p} \pmb{\textstyle\sum}_{p\times p} \textbf{V}_{p\times n}^{\pmb{T}} \\
        \pmb{K} =& \begin{bmatrix}
            u_0 & \hdots & u_{p-1}
        \end{bmatrix}
        \begin{bmatrix}
            \sigma_0&&\\
            &\ddots&\\
            &&\sigma_{p-1}
        \end{bmatrix}
        \begin{bmatrix}
            v_0 \\
            \vdots \\
            v_{p-1}
        \end{bmatrix}
\end{align*}
Note that it requires the number of rows and columns must be the same, we can also alleviate this by putting 0 appropriately,
to allow the kernel to be a square matrix.Then the first singular value $\sigma_0$ is checked. If it is the only one that is non zero,
and the rest is zero, then the kernel is separable. As an example, the sobel filter, which is used for edge detection, 
can be separated into two separate operations. In this case, this is for the horizontal derivative approximation. \cite{CHANG2023160}
\[
    \frac{1}{8}
    \begin{bmatrix}
        -1 & 0 & 1 \\
        -2 & 0 & 2 \\
        -1 & 0 & 1
    \end{bmatrix}
    =
    \frac{1}{8}
    \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
    \begin{bmatrix} 1 \\ 2 \\ 1 \end{bmatrix}
\]
The normalization of the operators are easily calculated by dividing each operator with the absolute sum of the entries.
\[  
    \begin{bmatrix} -1 & 0 & 1 \end{bmatrix} \Rightarrow \vert -1 \vert + \vert 0 \vert + \vert 1 \vert = 2 \Rightarrow \frac{1}{2}
\]
Then the normalized operator is
\[
    \frac{1}{2} \begin{bmatrix} -1 & 0 & 1 \end{bmatrix}
\]
it applies to the vertical operator as well
\[
    \frac{1}{4} \begin{bmatrix} 1 \\ 2 \\ 1 \end{bmatrix}
\]
The following are more examples of common separable filtering.

@@

\section{Simple examples of Linear Filters}
One notable example of a basic filter is the box filter, where each pixel is determined by the average of a $K \times K$ 
neighbordhood of pixels. It is also often called the \emph{moving average}, as the calculation can be done with scanlines, subtracting the
old pixel and adding the newest pixels to the running sum.
